---
title: "HW07"
author: "Zach White"
date: "10/28/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(MCMCpack)
library(MASS)
library(reshape2)
```

# Exercise 7.3

## Part A
```{r, cache = TRUE, warning = FALSE}
blue = read.table("/home/grad/zmw5/Fall 2016/STA 601/sta601/HW/bluecrab.dat")
orange = read.table("/home/grad/zmw5/Fall 2016/STA 601/sta601/HW/orangecrab.dat")
L.b.0 = S.b.0 = cov(blue)
L.o.0 = S.o.0 = cov(orange)
mu.b.0 = ybar.b = apply(blue,2,mean)
mu.o.0 = ybar.o = apply(orange,2,mean)

L.b.0.inv = S.b.0.inv = solve(L.b.0)
L.o.0.inv = S.o.0.inv = solve(L.o.0)

ybar.b = apply(blue,2,mean)
ybar.o = apply(orange,2,mean)
Sigma.b = cov(blue)
Sigma.o = cov(orange)
Sigma.b.inv = solve(Sigma.b)
Sigma.o.inv = solve(Sigma.o)

n.b = nrow(blue)
n.o = nrow(orange)
nu.0 = 4

n.iter = 10000

theta.b.post = theta.o.post = NULL
Sigma.b.post = Sigma.o.post = NULL




for(i in 1:n.iter){
  ### Update theta.b, theta.o
  Ln.b = solve(S.b.0.inv + n.b* Sigma.b.inv)
  Ln.o = solve(S.o.0.inv + n.o * Sigma.o.inv)
  mu.n.b = Ln.b %*% (L.b.0.inv %*% mu.b.0 + n.b*Sigma.b.inv %*% ybar.b)
  mu.n.o = Ln.o %*% (L.o.0.inv %*% mu.o.0 + n.o*Sigma.o.inv %*% ybar.o)
  theta.b = mvrnorm(1,mu.n.b,Ln.b)
  theta.o = mvrnorm(1,mu.n.o,Ln.o)
  
  theta.b.post = rbind(theta.b.post,theta.b)
  theta.o.post = rbind(theta.o.post,theta.o)
  
  ### Update Sigma.b, Sigma.o
  Sn.b = S.b.0 + (t(blue) - c(theta.b)) %*% t(t(blue) - c(theta.b))
  Sn.o = S.o.0 + (t(orange) - c(theta.o)) %*% t(t(orange) - c(theta.o))
  Sigma.b = solve(rWishart(1,nu.0 + n.b, solve(Sn.b))[,,1])
  Sigma.o = solve(rWishart(1,nu.0 + n.o, solve(Sn.o))[,,1])
  Sigma.b.post = rbind(Sigma.b.post,c(Sigma.b))
  Sigma.o.post = rbind(Sigma.o.post,c(Sigma.o))
}
```

## Part B
```{r, cache = TRUE, warning = FALSE}
theta.b.post = as.data.frame(theta.b.post)
names(theta.b.post) = c("theta1","theta2")

theta.o.post = as.data.frame(theta.o.post)
names(theta.o.post) = c("theta1","theta2")

par(mfrow = c(3,1))
plot(theta.b.post[,1], type = "l")
plot(theta.b.post[,2], type = "l")
plot(Sigma.b.post[,2] / sqrt(Sigma.b.post[,1]*Sigma.b.post[,4]), type = "l")

par(mfrow = c(3,1))
plot(theta.o.post[,1], type = "l")
plot(theta.o.post[,2], type = "l")
plot(Sigma.o.post[,2] / sqrt(Sigma.o.post[,1]*Sigma.o.post[,4]), type = "l")


both.theta.post = cbind(theta.b.post, theta.o.post)
id = rownames(both.theta.post)
both.theta.post = cbind(id,both.theta.post)
names(both.theta.post) = c("id","b.theta1","b.theta2","o.theta1","o.theta2")
both.theta.post = as.data.frame(both.theta.post)

par(mfrow = c(1,1))
ggplot(data = theta.b.post, aes(x = theta1, y = theta2)) + geom_point(color = "blue") + geom_point(data = theta.o.post, aes(x = theta1,y = theta2), color = "orange")

```

## Part C
```{r}
p.b = Sigma.b.post[,2] / sqrt(Sigma.b.post[,1] * Sigma.b.post[,4])
p.o = Sigma.o.post[,2] / sqrt(Sigma.o.post[,1]*Sigma.o.post[,4])

ggplot(data = as.data.frame(p.b) , aes(x = p.b)) + geom_density(color = "blue") + geom_density(data = as.data.frame(p.o), aes(x = p.o), color = "orange")

mean(p.b < p.o)
```


# Exercise 7.5
## Part A
```{r}
inter.exp = read.table("/home/grad/zmw5/Fall 2016/STA 601/sta601/HW/interexp.dat", header = TRUE)
apply(inter.exp,2,mean)
theta = sapply(inter.exp,mean,na.rm = TRUE)
thetaA = theta[1]
thetaB = theta[2]
## Calculate Sigma
sapply(inter.exp,var)
sigma2 = sapply(inter.exp,var,na.rm = TRUE)
sigma2.A = sigma2[1]
sigma2.B = sigma2[2]
## Calculate correlation matrix
complete.rho = cor(inter.exp, use = "complete")
rho = complete.rho[1,2]
```

## Part B
```{r}
A.missingB = inter.exp[is.na(inter.exp[,2]),1]
B.missingA = inter.exp[is.na(inter.exp[,1]),2]

impute.B = thetaB + (A.missingB - thetaA)* rho *sqrt(sigma2.B / sigma2.A)
impute.A = thetaA + (B.missingA - thetaB) * rho * sqrt(sigma2.A / sigma2.B)

imp.data = inter.exp
imp.data[is.na(imp.data[,2]),2] = impute.B
imp.data[is.na(imp.data[,1]),1] = impute.A

t.results = t.test(imp.data[,1],imp.data[,2], paired = TRUE)
t.results$conf.int
```

## Part C
I will use the unit information prior 
```{r, cache = TRUE}
ybar = apply(inter.exp,2,mean,na.rm = TRUE)
complete = which(complete.cases(inter.exp))
## Prior on Sigma
S = (t(inter.exp[complete,]) - ybar) %*% t(t(inter.exp[complete,]) - ybar)/length(complete)
nu.0 = nrow(S) + 2
n = nrow(inter.exp)

n.iter = 10000

y.A.samps = y.B.samps = matrix(0, nrow = n.iter,ncol = n)
theta.post = matrix(0,nrow=n.iter, ncol = 2)
names(theta.post) = c("thetaA","thetaB")

#STarting values
Y = imp.data
Sigma = S
theta = ybar

miss.A = which(is.na(inter.exp$yA))
miss.B = which(is.na(inter.exp$yA))

for(i in 1:n.iter){
  # Update theta
  y.bar.samp = apply(Y,2,mean)
  theta = mvrnorm(1,y.bar.samp,Sigma / (n+1))
  theta.post[i, ] <- theta
  
  # Update Sigma
  Sn<- S + ( t(Y)-c(theta) )%*%t( t(Y)-c(theta) )
  Sigma<-solve(rWishart(1, nu.0+n, solve(Sn))[, , 1])
  sigma2.A <- Sigma[1, 1]
  sigma2.B <- Sigma[2, 2]
  rho <- Sigma[1,2] / sqrt(sigma2.A*sigma2.B)
  
  for (j in miss.A) {
    Y[j, 1] <- rnorm(1, 
                     theta[1] + (rho*sqrt(sigma2.A/sigma2.B))*(Y[j, "yB"] - theta[2]),
                     sqrt4(sigma2.A*(1 - rho)))
  }
  for (j in miss.B) {
    Y[j, 2] <- rnorm(1, 
                     theta[2] + (rho*sqrt(sigma2.B/sigma2.A))*(Y[j, "yA"] - theta[1]),
                     sqrt(sigma2.B*(1 - rho)))
  }
  
  
  #Y[miss.A,1] = rnorm(length(miss.A),(rho*sqrt(sigma2.A/sigma2.B))*(Y[miss.A,"yB"] - theta[2]),sqrt(sigma2.A*(1-rho)))
  #Y[miss.B,2] = rnorm(length(miss.B),(rho*sqrt(sigma2.B/sigma2.A))*(Y[miss.B,"yA"] - theta[1]), sqrt(sigma2.B*(1-rho)))
  
  y.A.samps[i, ] <- Y[, 1]
  y.B.samps[i, ] <- Y[, 2]
}

theta.diff = theta.post[,1] - theta.post[,2] 
cred.int = quantile(theta.diff, c(.025,.975))
cred.int
```

The credible interval of interest is `r cred.int`, which means that there is a .95 probability that $\theta_A-\theta_B | y$ lies in th interval `r cred.int`.  When comparing this result with the results from the t test, ours are a little bit wider.  Our lower extreme is lower than the one from the traditional t test.  However, we now have significantly more information with our analysis because we have posteriors on $\Sigma, \theta_A,\theta_B$ and finally we actually have distributions on the imputed values.

# Exercise 8.2

$$
\begin{aligned}
\mu &\sim N(\mu_0,\gamma_o^2) = N(75,100) \\
\ \frac{1}{\sigma^2} &\sim Gamma(\eta_0,\tau_0^2) = Gamma(1,100) \\
\ \delta &\sim N(\delta_0,\tau_0^2) \text{ where} \\
\delta_0 &\in \{-4,-2,0,2,4\} \text{ & } \tau_0^2 \in \{10,50,100,500\}

\end{aligned}
$$

## Part A
```{r, cache = TRUE}
n.a = n.b = 16
y.bar.a = 75.2
s.a = 7.3
y.bar.b = 77.5
s.b = 8.1
sum.y.a = n.a * y.bar.a
sum.y.b = n.b * y.bar.b

delta0 = seq(-4,4,by = 2)
tau2.0 = c(10,50,100,500)



# Hyperparams
mu.0 = 75
gamma2.0 = 100
nu.0 = 1
sigma2.0 = 100


# Starting values
mu = (y.bar.a + y.bar.b )/2
delta = (y.bar.a - y.bar.b) / 2

# Form some arays to store values

mu.array =sigma2.array = delta.array =  array(0,c(length(delta0),length(tau2.0),n.iter))

  
## Gibbs Sampler
n.iter = 10000
for(i in 1:n.iter){
  for(j in 1:length(delta0)){
    for(k in 1:length(tau2.0)){
  
      # Update Sigma2
      nu.sigma = (nu.0*sigma2.0 + (n.a -1)*s.a^2 + n.a*(y.bar.a - (mu + delta))^2 + (n.b-1)*s.b^2 + n.b*(y.bar.b - (mu-delta))^2)
      sigma2 = 1 / rgamma(1,(nu.0 + n.a + n.b)/2,nu.sigma)
      sigma2.array[j,k,i] = sigma2
      
      ## Update mu
      var.mu = 1/(1/gamma2.0 + (n.a + n.b)/sigma2)
      mean.mu = var.mu * (mu.0 / gamma2.0 +  ((sum.y.a - n.a*mu)/sigma2) + ((sum.y.b + n.b*mu)/sigma2))
      mu = rnorm(1,mean.mu,sqrt(var.mu))
      mu.array[j,k,i] = mu
      
      ## Update delta
      var.delta = 1/(1/tau2.0[k] + (n.a + n.b)/sigma2)
      mean.delta = var.delta*(delta0[j]/tau2.0[k] + ((sum.y.a - n.a*mu)/sigma2) - (sum.y.b - n.b*mu)/sigma2)
      delta = rnorm(1,mean.delta,sqrt(var.delta))
      delta.array[j,k,i] = delta
    }
  }
}

## Post Analysis
#sapply(delta.array,c(1,2), mean < 0)
prob.0 = lower.bound = upper.bound = post.cor = matrix(0,5,4)
delta.n = length(delta0)
tau.n = length(tau2.0)
theta.A = mu.array + delta.array
theta.B = mu.array - delta.array
for(i in 1:delta.n){
  for(j in 1:tau.n){
    prob.0[i,j] = mean(delta.array[i,j,] < 0)
    lower.bound[i,j] = quantile(delta.array[i,j,],.025)
    upper.bound[i,j] = quantile(delta.array[i,j,],.975)
    post.cor[i,j] = cor(theta.A[i,j,],theta.B[i,j,])
  }
}
prob.0
apply(delta.array,c(1,2),quantile,c(.025,.975))
apply(delta.array,c(1,2),quantile,.025)
apply(delta.array,c(1,2),quantile,.975)
```

This matrix shows `r prob.0` for the varying combinations of the prior specifications for $\delta_0$ and $\tau_0^2$.

```{r}
lower.bound
upper.bound
```

```{r}
theta.A = mu.array + delta.array
theta.B = mu.array - delta.array
```

## Part B

